\Large{ Authors' reply to the comments by Anonymous Referee (Referee \#2) of the manuscript gmd-2017-119} \\ \LARGE \textbf{Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed particle hydrodynamics}

### 2018-03-17 Manuscripts_GMD_ap_ref2.tex
updated the manuscripts

The Plume-SPH code also incorporates  several newly developed techniques in SPH  needed to address numerical challenges in simulating multiphase compressible turbulent flow.  The code should thus be also of general interest to the much larger community of researchers using and developing SPH based tools. In particular, the $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales, heat exchange due to turbulence is calculated by a Reynolds analogy and a corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries. We also developed methodology to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.

@@

The development of 2D and 3D, time-dependent, and multiphase numerical models for volcanic plumes has provided new explanations for many features of explosive volcanism. One of the earliest of these is the 3D model PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model. Conservation equations for each phase are solved separately with the finite volume method. A parallel computing version of PDAC was also developed \citep{ongaro2007parallel}. Advanced numerical techniques, such as a second order scheme and semi-explicit time stepping, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.

@@

\subsection{Governing equations}
Based on above assumptions, the governing equations of our model are given as (which is the same as the governing equations of SK-3D \citep{suzuki2005numerical}):
\begin{align}
\dfrac{\partial \rho}{\partial t} + \nabla \cdot \left(\rho \textbf{v}\right) = 0 \label{eq:gov-cs-rho} \\
\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot \left(\rho \xi \textbf{v}\right) = 0 \label{eq:gov-cs-ks}\\
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot \left(\rho \textbf{v} \textbf{v} + p\textbf{I}\right) = \rho \textbf{g} \label{eq:gov-cs-v} \\
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot \left[\left(\rho E + p \right)\textbf{v}\right] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
\end{align}
where $\rho$ is the density, $\textbf{v}$ is the velocity, $\xi$ is the mass fraction of ejected material, $\textbf{g}$ is the gravitational acceleration, $\textbf{I}$ is a unit tensor.
$E = e + K $ is the total energy which is a summation of kinetic energy $K$ and internal energy $e$.
An additional equation is required to close the system. In this model, the equation for closing the system is the following EOS (equation of state).
\begin{equation}
p = \left(\gamma_m - 1\right)\rho e \label{eq:EOS}
\end{equation}
where 
\begin{equation}
\gamma_m = R_m/C_{vm} + 1 \label{eq:gov-gm}
\end{equation}
\begin{equation}
R_m = \xi_g R_g + \xi_a R_a  \label{eq:gov-Rm}
\end{equation}
\begin{equation}
C_{vm} = \xi_s C_{vs} + \xi_g C_{vg} + \xi_a C_{va} \label{eq:gov-Cvm}
\end{equation}
\begin{equation}
\xi_a = 1 - \xi \label{eq:gov-na}
\end{equation}
\begin{equation}
\xi_g = \xi \cdot \xi_{g0} \label{eq:gov-ng}
\end{equation}
\begin{equation}
\xi_s = \xi - \xi_g \label{eq:gov-ns}
\end{equation}
where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant. $\xi$ with subscript is the mass fraction of corresponding constituent. The subscript 
$m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material and $a$ represents air. $\xi_{g0}$ is the mass fraction of vapor in the erupted material.

@@

\subsubsection{Velocity inlet}
At the vent, temperature of erupted material $T$, eruption velocity $\textbf{v}$, the mass fraction of vapor in erupted material $\xi_{g0}$ and mass discharge rate $\dot M$ are given. The pressure of erupted material $p$ is assumed to be the same as ambient pressure for pressure-balanced eruption. The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$. Equation (\ref{eq:erupt_bc_rho}) to (\ref{eq:erupt_bc_e}) gives velocity inlet boundary condition wrote in terms of primitive variables.
\begin{align}
\rho =const = p/\left(R_m T\right) \label{eq:erupt_bc_rho} \\
\xi=const=1 \label{eq:erupt_bc_xi}\\
\textbf{v} = const =\{u,v,w\}^T \label{eq:erupt_bc_v}\\
\dfrac{\partial e}{\partial n}=\dot M e /\left(\pi r^2\right) \label{eq:erupt_bc_e}
\end{align} 
where $r$ is the radius of the vent, $n$ is the direction normal to the boundary.

@@

\subsection{Mass fraction update}
Air and erupted material are represented by two different sets of SPH particles (or discretization points) in the model. Based on assumptions we made in section \ref{sec:physics-model}, only density needs to be updated respectively for each phase. The updating of density is exactly the same as Eq. (\ref{eq:ns-sph-d}) in spirit.  Particles of phase 1 are not counted while evaluating density of phase 2 and vice versa. Updating of density is then based on the following discretized equations.
\begin{equation}
<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d1}
\end{equation}
\begin{equation}
<\rho_\alpha^{sg}>=\frac{\sum m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d2}
\end{equation}
where the subscript $a$ and $b$ represents air particles (phase 1) while $i$ and $j$ represents erupted material. $\beta$ and $\alpha$ represents either erupted material or air.
$\rho_a^a$ is density of phase 1 (air). 
 $\rho_i^{sg}$ is density of phase 2 (erupted material).
$\rho=\rho^a + \rho^{sg}$ is density of mixture of air and erupted material. 
By definition, the mass fraction is updated according to Eq. (\ref{eq:gov-sph-xi}).
\begin{equation}
<\xi_{\alpha}> = \dfrac{\rho^{sg}_{\alpha}}{\rho_{\alpha}}
\label{eq:gov-sph-xi}
\end{equation}

@@

Interface construction will become necessary and important when attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence. 
As a Lagrangian method, interface tracking in SPH is explicit through capturing of the locations of the particles, much simplier than Euleruan Methods.
The existence of complex evolving interfaces between phases presents severe challenges to conventional Eulerian grid-based numerical methods. Either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow. High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of tracking techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}. For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which needs additional numerical effort \citep{hirt1981volume, youngs1982time}.
Since SPH is able to adaptively adjust the discretization and automatically construct the interface. SPH requires less additional numerical effort for interface construction and therefore is more suitable for volcanic plume simulation.

@@

\subsection{Turbulence modeling with SPH}
For high speed shearing flow, the momentum exchange and heat transfer are dominated by turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical coefficients (molecular viscosity and heat conduction coefficient). In addition to momentum and enrgy exchange, mixing between plume and air is important in plume modeling. Quantifying these mixing processes in real implementation is challenging because of the scale disparity between the large-scale fluid motion and the diffusion processes on interface that ultimately lead to mixing. 
%The energy transfer between these scales occurs through turbulent motion, created either by fluid instabilities or by breaking internal waves. For volcanic plume, a shearing flow, the fluid instabilities dominated the mixing process. 
One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) turbulence model. 
Another way is to use fine enough resolution (namely, the direct numerical simulation method) at the expense of much higher computational cost which results from both increase in number of particles and decrease in time steps constrained by the CFL (Courant-Friedrichs-Lewy) condition. Here we choose the SPH-SPS method. Among existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}. However, the $SPH-\varepsilon$ turbulence model was proposed only for incompressible flow. In the following section, we will extend it for compressible flow. It is necessary to mention that all other existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical} also only focus on incompressible flow.

@@

\subsection{Boundary conditions}
All boundary conditions are imposed by ghost particles. Subfigure $b$ of Fig. \ref{fig:bc_and_domain_decomp} shows how boundaries are deployed. 
%Physical quantities of particles, including real particles and different types of ghost particles, are updated differently.
\subsubsection{Wall boundary condition}
Traditionally either ghost particles that mirror real particles across the boundary \citep {ferrari2009new} or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions. One disadvantage of the latter  is that the boundary forces tend to corrupt the solution in the local neighborhood. In addition, a natural way of imposing eruption boundary condition is using eruption ghost particles. To impose boundary conditions in a consistent way, we adopt a modified version of the ghost particle method \citep {kumar2013parallel} for wall boundary conditions. Stationary wall ghost particles are deployed in the same way as real particles. Instead of enforcing symmetry particle by particle, a symmetric field across the boundary is explicitly enforced. Ghost particles are reflected into the domain and physical quantities are calculated at these reflected positions by SPH interpolations. It should be noted that wall ghost particles should not be counted when computing physical properties of these reflected positions. Assign all properties, except for velocity, at the corresponding reflected position to the ghost particle. The velocity of each wall ghost particles is set to have the same value but opposite direction of the interpolated velocity at its corresponding reflection. By this way, the no-slip wall boundary condition (Eq. (\ref{eq:wall_bc_v})) is imposed naturally. These wall ghost particles serve as neighbors in momentum and energy update. More implementation details about this method can be found in \citep {kumar2013parallel}. As these wall ghost particles are stationary, there is no mass flux on the boundary (Eq. (\ref{eq:wall_bc_rho}) and (\ref{eq:wall_bc_xi})). In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature vanishes and hence there is no internal energy flux on the wall boundary (Eq. (\ref{eq:wall_bc_e})). 
In our current model, the ground is assumed to be  flat. For more complicated topography, it has been shown in other work \citep {kumar2013parallel} that it also works as well. We do have some concern regarding potential limitation of this method of deployment of ghost particles for more complicated boundaries in three dimensions but that is one that is not related to the plume simulations that are the target here.

@@

Any implementation of SPH code requires efficient searching and updating of neighbors during simulation. Of the many possible choices we adopt a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH. This background grid is also used for domain decomposition in SPH. We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimensions, as buckets. 
As for the actual storage of the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.

@@

As for domain decomposition, even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation. So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient}. Subfigure $c$ of Fig. \ref{fig:bc_and_domain_decomp} shows a cross section view of domain decomposition.

@@

%\begin{figure}[!t]
%\centering
%\includegraphics[scale=0.35]{strong}
%\caption{Strong scalability test result}
%\label{fig:strong_scale}
%\end{figure}
%
%\begin{figure}[!t]
%\centering
%\includegraphics[scale=0.35]{weak_scale}
%\caption{Weak scalability test result}
%\label{fig:weak_scale}
%\end{figure}
%
\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{adj_vs_no}
\caption{The effect of domain adjusting on simulation time. The figure on the top shows execution time without domain adjusting, the figure on the bottom shows execution time with domain adjusting. Different bins represent execution time up to specific physical time indicated by horizontal axis.}
\label{fig:adj_vs_no}
\end{figure}

@@

\begin{figure}[!htb]
    \centering
    \begin{minipage}{.45\textwidth}
        \centering
        \includegraphics[width=0.99 \textwidth]{./mssfrc-Diverging}
    \end{minipage}%
    \begin{minipage}{.45 \textwidth}
        \centering
        \includegraphics[width=0.99 \textwidth]{./mssfrc-Diverging-cut}
    \end{minipage}%
    \\
    \begin{minipage}{.45 \textwidth}
        \centering
        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-yz}
    \end{minipage}% 
    \begin{minipage}{.45 \textwidth}
        \centering
        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-yz-zoomed}
    \end{minipage}%  
    \caption{Mass fraction for $t=300s$ after eruption. Figures on the first row are visualization of SPH simulation results. The figure on the up right corner is visualization of a slice of the computational, whose thickness is around $10000m$. The lowest portion of the plume represents erupted material in eruption vent (the underground portition). Figures at the second row show contour of mass fraction and velocity quiver on y-z plane. These figures are plotted utilizing post-processed data (see Appendix \ref{app:project-SPH-grid}). The contour level in the plot are $0.00001, 0.0001, 0.001, 0.1, 0.3, 0.75, 0.95$. The figure on the lower right corner is a zoomed view of velocity quivier showing plume expansion and entrainment of air.}
    \label{fig:pinatubo-simulation-results-vis}
\end{figure}

@@

\conclusions  \label{sec:conclusion}%% \conclusions[modified heading if necessary]
A new plume model was developed based on the SPH method. Extensions necessary for Lagrangian methodology and compressible flow were made in the formulation of the equations of motion and turbulence models. Advanced numerical techniques in SPH were exploited and tailored for this model. High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading) and simulation time (depends on comprehensiveness of model, resolution, order of accuracy of numerical methods, scheme for time upgrading ... and computational techniques). The correctness of the code and model was verified and validated by a series of test simulations. Typical 1D shock tube problems were simulated and compared against analytical results showing good agreement. Dimensionless velocity and concentration distribution across the cross-section and along the jet axis match well with experimental results of JPUE. Top height and integrated local variables simulated by our model are consistent with simulation results of existing 3D plume models. Comparison between our results with those of SK-3D implies that turbulence model plays a significant role in plume modeling.

@@

We have presented in this paper an initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and high performance computing. More advanced numerical techniques, such as adaptive particle size, Godunov-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit in the future. In the near future, effect of wind field will be take into account. Our code will also be made available in the open source form for the community to enhance.
%In all existing 3D plume models, numerical calculations are performed on a non-uniform (vertically and horizontally stretched) grid, with different grid resolutions and domain sizes. However, implementation of different resolution in SPH methods is not straight forward and requires more work.

>>>>

The Plume-SPH code also incorporates  several newly developed techniques in SPH  needed to address numerical challenges in simulating multiphase compressible turbulent flow. The code should thus be also of general interest to the much larger community of researchers using and developing SPH based tools. In particular, the $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales. Heat exchange due to turbulence is calculated by a Reynolds analogy and a corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries. We also developed methodology to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.

@@

The development of 2D and 3D, time-dependent, and multiphase numerical models for volcanic plumes has provided new explanations for many features of explosive volcanism. For example, a recent study based on 3D fluid dynamical model \citep{costa2018understanding} shows that simple extrapolations of integral models for Plinian columns to those of super-eruption plumes are not valid and their dynamics diverge from current ideas of how volcanic plumes operate. One of the earliest of these is the 3D model PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model. Conservation equations for each phase are solved separately with the finite volume method. A parallel computing version of PDAC was also developed \citep{ongaro2007parallel}. Advanced numerical techniques, such as a second order scheme and semi-explicit time stepping, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.

@@

\subsection{Governing equations}
Based on above assumptions, the governing equations of our model are given as (which is the same as the governing equations of SK-3D \citep{suzuki2005numerical}):
\begin{align}
\dfrac{\partial \rho}{\partial t} + \nabla \cdot \left(\rho \textbf{v}\right) = 0 \label{eq:gov-cs-rho} \\
\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot \left(\rho \xi \textbf{v}\right) = 0 \label{eq:gov-cs-ks}\\
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot \left(\rho \textbf{v} \textbf{v} + p\textbf{I}\right) = \rho \textbf{g} \label{eq:gov-cs-v} \\
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot \left[\left(\rho E + p \right)\textbf{v}\right] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
\end{align}
where $\rho$ is the density, $\textbf{v}$ is the velocity, $\xi$ is the mass fraction of ejected material, $\textbf{g}$ is the gravitational acceleration, $\textbf{I}$ is a unit tensor.
$E = e + K $ is the total energy which is a summation of kinetic energy $K$ and internal energy $e$.
An additional equation is required to close the system. In this model, the equation for closing the system is the following EOS (equation of state).
\begin{equation}
p = \left(\gamma_m - 1\right)\rho e \label{eq:EOS}
\end{equation}
where 
\begin{equation}
\gamma_m = R_m/C_{vm} + 1 \label{eq:gov-gm}
\end{equation}
\begin{equation}
R_m = \xi_g R_g + \xi_a R_a  \label{eq:gov-Rm}
\end{equation}
\begin{equation}
C_{vm} = \xi_s C_{vs} + \xi_g C_{vg} + \xi_a C_{va} \label{eq:gov-Cvm}
\end{equation}
\begin{equation}
\xi_a = 1 - \xi \label{eq:gov-na}
\end{equation}
\begin{equation}
\xi_g = \xi \cdot \xi_{g0} \label{eq:gov-ng}
\end{equation}
\begin{equation}
\xi_s = \xi - \xi_g \label{eq:gov-ns}
\end{equation}
where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant. $\xi$ is the mass fraction of. The subscript $m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material, $a$ represents air, $0$ represents phyiscal properties of erupted material. $\xi_{g0}$ is the mass fraction of vapor in the erupted material.

@@

\subsubsection{Velocity inlet}
At the vent, temperature of erupted material $T$, eruption velocity $\textbf{v}$, the mass fraction of vapor in erupted material $\xi_{g0}$ and mass discharge rate $\dot M$ are given. The pressure of erupted material $p$ is assumed to be the same as ambient pressure for pressure-balanced eruption. The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$. Equation (\ref{eq:erupt_bc_rho}) to (\ref{eq:erupt_bc_e}) gives velocity inlet boundary condition wrote in terms of primitive variables.
\begin{align}
\rho =const = p/\left(R_m T\right) \label{eq:erupt_bc_rho} \\
\xi=const=1 \label{eq:erupt_bc_xi}\\
\textbf{v} = const =\{u,v,w\}^T \label{eq:erupt_bc_v}\\
\dfrac{\partial e}{\partial n}=\dot M e_0 /\left(\pi r^2\right) \label{eq:erupt_bc_e}
\end{align} 
where $r$ is the radius of the vent, $n$ is the direction normal to the boundary, $e_0$ is internal energy of erupted material which can be calculated based on temperature and specific heat of erupted material.

@@

\subsection{Mass fraction update}
Air and erupted material are represented by two different sets of SPH particles (or discretization points) in the model. Based on assumptions we made in section \ref{sec:physics-model}, only density needs to be updated respectively for each phase. The updating of density is exactly the same as Eq. (\ref{eq:ns-sph-d}) in spirit.  Particles of phase 1 are not counted while evaluating density of phase 2 and vice versa. Updating of density is then based on the following discretized equations.
\begin{equation}
<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d1}
\end{equation}
\begin{equation}
<\rho_\alpha^{sg}>=\frac{\sum m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d2}
\end{equation}
where the subscript $a$ and $b$ represents air particles (phase 1) while $i$ and $j$ represents particles of erupted material. $\beta$ and $\alpha$ represents either erupted material particles or air particles.
$\rho_a^a$ is density of phase 1 (air). 
 $\rho_i^{sg}$ is density of phase 2 (erupted material).
$\rho=\rho^a + \rho^{sg}$ is density of mixture of air and erupted material. 
By definition, the mass fraction is updated according to Eq. (\ref{eq:gov-sph-xi}).
\begin{equation}
<\xi_{\alpha}> = \dfrac{\rho^{sg}_{\alpha}}{\rho_{\alpha}}
\label{eq:gov-sph-xi}
\end{equation}

@@

Interface construction will become necessary and important when attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence. 
As a Lagrangian method, interface tracking in SPH is explicit through capturing of the locations of the particles, much simplier than Euleruan methods.
The existence of complex evolving interfaces between phases presents severe challenges to conventional Eulerian grid-based numerical methods. Either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow. High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of tracking techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}. For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which needs additional numerical effort \citep{hirt1981volume, youngs1982time}.
Since SPH is able to adaptively adjust the discretization and automatically construct the interface. SPH requires less additional numerical effort for interface construction and therefore is more suitable for volcanic plume simulation.

@@

\subsection{Turbulence modeling with SPH}
For high speed shearing flow, the momentum exchange and heat transfer are dominated by turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical coefficients (molecular viscosity and heat conduction coefficient). In addition to momentum and enrgy exchange, mixing between plume and air is important in plume modeling. Quantifying these mixing processes in real implementation is challenging because of the scale disparity between the large-scale fluid motion and the diffusion processes on interface that ultimately lead to mixing. Ideally, one would like to be able to include the effects of mixing on the large scale dynamics without resolving the detailed interface structure and dynamics of turbulence to reduce computational cost.
To resolve all turbulent exchange at all different scales and sub-particle scale mixing with relative coarse resolution, a SPH-SPS turbulence models should be included. Among existing SPH-SPS (SPH sub-particle-scale) turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}. However, the $SPH-\varepsilon$ turbulence model was proposed only for incompressible flow. In the following section, we will extend it for compressible flow. It is necessary to mention that all other existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical} also only focus on incompressible flow.

@@

\subsection{Boundary conditions}
All boundary conditions are imposed by ghost particles. Subfigure $b$ of Fig. \ref{fig:bc_and_domain_decomp} shows how boundaries are deployed. 
%Physical quantities of particles, including real particles and different types of ghost particles, are updated differently.
\subsubsection{Wall boundary condition}
Traditionally either ghost particles that mirror real particles across the boundary \citep {ferrari2009new} or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions. One disadvantage of the latter  is that the boundary forces tend to corrupt the solution in the local neighborhood. In addition, a natural way of imposing eruption boundary condition is using eruption ghost particles. To impose boundary conditions in a consistent way, we adopt a modified version of the ghost particle method \citep {kumar2013parallel} for wall boundary conditions. Stationary wall ghost particles are deployed in the same way as real particles. Instead of enforcing symmetry particle by particle, a symmetric field across the boundary is explicitly enforced. Ghost particles are reflected into the domain and physical quantities are calculated at these reflected positions by SPH interpolations. It should be noted that wall ghost particles should not be counted when computing physical properties of these reflected positions. Assign all properties, except for velocity, at the corresponding reflected position to the ghost particle. The velocity of each wall ghost particles is set to have the same value but opposite direction of the interpolated velocity at its corresponding reflection. By this way, the no-slip wall boundary condition (Eq. (\ref{eq:wall_bc_v})) is imposed naturally. These wall ghost particles serve as neighbors in momentum and energy update. More implementation details about this method can be found in \citep {kumar2013parallel}. As these wall ghost particles are stationary, there is no mass flux on the boundary (Eq. (\ref{eq:wall_bc_rho}) and (\ref{eq:wall_bc_xi})). In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature vanishes and hence there is no internal energy flux on the wall boundary (Eq. (\ref{eq:wall_bc_e})). 
In our current model, the ground is assumed to be  flat. For more complicated topography, it has been shown in other work \citep {kumar2013parallel} that it also works as well. We do have concern regarding potential limitation of this method of deployment of ghost particles for more complicated boundaries in three dimensions. But that is not related to the plume simulations in current plume model.

@@

The time complexity for raw SPH method is $O(NN)$. Where, $N$ is the total number of particles (SPH particles are essentially discretization points and equivalent to grids in mesh-based methods). Efficient neighbour searching and compact-supported kernel function helps reduce computational cost. Of the many possible choices we adopt a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH. Then the time complexity is reduced to $O(MN+mN)$. Where $m$ is the average of number of particles within the compact support of kernel function, $M$ is average number of particles in sub-domain within which neighbour searching is carried out. This background grid is also used for domain decomposition. We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimensions, as buckets. 
As for the actual storage of the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.

@@

The parallelization is achieved by splitting computational domain into subdomains. Each subdomain is computed by a single processor. For any subdomain, information from its neighbouring subdomains is required when updating physical quantities. To guarantee consistency, data is synchronized after each updating of physical quantity. Even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation. So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient} to decompose the computational domain. Subfigure $c$ of Fig. \ref{fig:bc_and_domain_decomp} shows a cross section view of domain decomposition.

@@

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{adj_vs_no}
\caption{The effect of domain adjusting on simulation time. The figure on the top shows execution time without domain adjusting, the figure on the bottom shows execution time with domain adjusting. Different bins represent execution time up to specific physical time indicated by horizontal axis.}
\label{fig:adj_vs_no}
\end{figure}

@@

\begin{figure}[!htb]
    \centering
    \begin{minipage}{.45\textwidth}
        \centering
        \includegraphics[width=0.99 \textwidth]{./mssfrc-Diverging}
    \end{minipage}%
    \begin{minipage}{.45 \textwidth}
        \centering
        \includegraphics[width=0.99 \textwidth]{./mssfrc-Diverging-cut}
    \end{minipage}%
    \\
    \begin{minipage}{.45 \textwidth}
        \centering
        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-xz}
    \end{minipage}% 
    \begin{minipage}{.45 \textwidth}
        \centering
        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-xz-zoomed}
    \end{minipage}%  
    \caption{Mass fraction for $t=500s$ after eruption. Figures on the first row are visualization of SPH simulation results. The figure on the up right corner is visualization of a slice of the computational, whose thickness is around $10000m$. The lowest portion of the plume represents erupted material in eruption vent (the underground portition). Figures at the second row show contour of mass fraction and velocity quiver on x-z plane. These figures are plotted utilizing post-processed data (see Appendix \ref{app:project-SPH-grid}). The contour level in the plot are $0.00001, 0.0001, 0.001, 0.1, 0.3, 0.75, 0.95$. The figure on the lower right corner is a zoomed view of velocity quivier showing plume expansion and entrainment of air.}
    \label{fig:pinatubo-simulation-results-vis}
\end{figure}

@@

\conclusions  \label{sec:conclusion}%% 
A new plume model was developed based on the SPH method. Extensions necessary for Lagrangian methodology and compressible flow were made in the formulation of the equations of motion and turbulence models. Advanced numerical techniques in SPH were exploited and tailored for this model. High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading) and simulation time (depends on comprehensiveness of model, resolution, order of accuracy of numerical methods, scheme for time upgrading ... and computational techniques). The correctness of the code and model was verified and validated by a series of test simulations. Typical 1D shock tube problems were simulated and compared against analytical results showing good agreement. Dimensionless velocity and concentration distribution across the cross-section and along the jet axis match well with experimental results of JPUE. Top height and integrated local variables simulated by our model are consistent with simulation results of existing 3D plume models. Comparison between our results with those of SK-3D implies that turbulence model plays a significant role in plume modeling.

@@

We have presented in this paper an initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and high performance computing. More advanced numerical techniques, such as adaptive particle size, Godunov-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit in the future. In the near future, effect of wind field will be take into account. Our code will also be made available in the open source form for the community to enhance. Besides improving plume model, coupling volcanic plume model with magma reservoir models \citep[e.g.][]{terray2018new}, which could provide more accurate eruption conditions, might improve accuracy of volcanic plume simulation.

### 2018-03-17 Reply-to-RC2.tex
updated the manuscripts

Without accounting for detailed algorithm, the time complexity for raw SPH method is $O(NN)$. Where, $N$ is the total number of particles (SPH particles are essentially discretization points and equivalent to grids in mesh-based methods). By adopting background mesh and compact-supported kernel function, the time complexity will be reduced to $O(MN)$. Where $M$ is the average of number of particles within the compact support of kernel function. Please be noted that number of particles within the compact support of kernel function is not constant in our case.

When computations are performed across different processors, theoretically speaking, the total computational time will be $(s+\frac{p}{n}+\bar{p})$. Where $(s+p)\sim O(MN)$ is sequential computational time with $p$ representing computation that parallelizable and $s$ representing computation that can not be parallelized. $n$ is total number of processors. $\bar{p}$ is extra computation introduced by parallelization, for example, 
unecessary duplicating work, communication overhead, time to split and combine.

>>>>

Without accounting for detailed algorithm, the time complexity for raw SPH method is $O(NN)$. Where, $N$ is the total number of particles (SPH particles are essentially discretization points and equivalent to grids in mesh-based methods). By adopting background mesh and compact-supported kernel function, the time complexity will be reduced to $O(MN+mN)$. Where $m$ is the average of number of particles within the compact support of kernel function, $M$ is number of particles among which neighbour searching is carried out. Please be noted that number of particles within the compact support of kernel function is not constant in our case.

When computations are performed across different processors, theoretically speaking, the total computational time will be $(s+\frac{p}{n}+\bar{p})$. Where $(s+p)\sim O(MN+mN)$ is sequential computational time with $p$ representing computation that parallelizable and $s$ representing computation that can not be parallelized. $n$ is total number of processors. $\bar{p}$ is extra computation introduced by parallelization, for example, 
unecessary duplicating work, communication overhead, time to split and combine.
